---
title: "covid_analysis"
output: html_document
date: "2023-09-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(writexl)
library(ggplot2)
library(janitor)
library(readxl)
#install.packages("plyr")
library(plyr)
```
# victoria adding in and cleaning new covid historic data 9-20-2022 - 9-05-2023
```{r}
# read in covid positive report data 9-20-2022 - 9-05-2023
data23 <- read_csv('newdatarightformat.csv')

#get rid of totals rows for each year
data23 <- data23[-c(1,105,293),]

# clean names and create 1 column for role type instead of having 1 column for each different role
data23 <- data23 %>%
  clean_names() %>%
  pivot_longer(cols = c(contains('faculty'), contains('staff'), contains('student')))

# rename the columns to match what the old dataframe below has
colnames(data23) <- c('date','drop','role','count')

# make the dates consistent with the old datas format
data23$date <- as.Date(parse_date_time(data23$date, "mdy"))

# select only the columns I need, we don't need the column that has the totals for each day
data23 <- data23 %>%
  select(role, date, count)

# make all the NAs 
data23[is.na(data23)] <- 0

```
# this is Devon's old code, I have commented where I adjusted it
```{r}
# this is all devon adding in the dataframe we had last year, adjusting column names, getting rid of nas

data <- read_csv('data930.csv')
colnames(data) <- c('drop','type','role','date','count')
data <- data%>%
drop_na(drop)
data <- subset(data, select = -c(drop))

#victoria 9/21/2023 also dropping type column from the old data to match our current data

data <- subset(data, select = -c(type))

# this was devon, just adjustng the date
data$date <- as.Date(parse_date_time(data$date, "dmy"))

#victoria 9/21/2023 making all nas equal to zero in old data

data[is.na(data)] <- 0

```

#victoria 9/21/2023 joining the dataframe 'data' and the dataframe 'data23' to create a dataframe for total historic recorded positive covid cases

```{r}
# joining the two dataframes
historic_data <- rbind(data, data23) %>%
  # removing the role column because we really just want the totals 
  select(-c(role)) %>%
  
# now I'm summing the case count for each date  
historic_data <- aggregate(.~date,data=historic_data,FUN=sum)

historic_data <- historic_data %>%
  mutate(sum = rowSums(across(where(is.numeric)), na.rm=TRUE))

# and then I'm writing it out to this file 

write_xlsx(historic_data, "newcasedata.xlsx")

```

# Victoria - now I'm using the new complete dataframe to see how many cases were in the same time period (first 9 days of the semester) this year compared to last year

```{r}

v_fall2022  <- historic_data[historic_data$date >= "2022-08-29" & historic_data$date <= "2022-09-06",]

v_fall2023  <- historic_data[historic_data$date >= "2023-08-28" & historic_data$date <= "2023-09-05",]


sum(v_fall2022$count)
sum(v_fall2023$count)

# This is the end of victoria stuff

```



```{r}
#create dfs by semester
fall2020 <- historic_data[historic_data$date >= "2020-08-30" & historic_data$date <= "2020-12-22", ]
spring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-05-19", ]
fall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-12-22", ]
spring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-05-18", ]
fall2022  <- historic_data[historic_data$date >= "2022-08-29",]


#devon semesters - adjust start date for fall 2020, end date for fall 2021 -- USE THIS FOR TOTAL SEM COUNT
d_fall2020 <- historic_data[historic_data$date >= "2020-08-31" & historic_data$date <= "2020-12-22", ]
d_spring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-05-19", ]
d_fall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-12-21", ]
d_spring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-05-18", ]
d_fall2022  <- historic_data[historic_data$date >= "2022-08-29" & historic_data$date <= "",]

```

```{r}
sum(fall2020$count)
sum(spring2021$count)
sum(fall2021$count)
sum(spring2022$count)
sum(fall2022$count)

#devon sums with new adjusted semesters -- USE THIS FOR TOTAL SEM COUNT
sum(d_fall2020$count)
sum(d_spring2021$count)
sum(d_fall2021$count)
sum(d_spring2022$count)
sum(d_fall2022$count)



```
#3 dont use this block
```{r}
shortfall2020 <- historic_data[historic_data$date >= "2020-08-30" & historic_data$date <= "2020-09-21", ]
shortspring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-02-10", ] #why is this shorter? - devon
shortfall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-09-21", ]
shortspring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-02-11", ] #why is this shorter? - devon
fall2022  <- historic_data[historic_data$date >= "2022-08-29",]

#devon short sems adjusting fall 2020 start, spring 2021 end, fall 2021 end, spring 2022 end
d_shortfall2020 <- historic_data[historic_data$date >= "2020-08-31" & historic_data$date <= "2020-09-21", ]
d_shortspring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-02-15", ]
d_shortfall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-09-20", ]
d_shortspring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-02-14", ]
d_fall2022  <- historic_data[historic_data$date >= "2022-08-29",]
```

## good block
```{r}

#again, again with new historic_data --> 32 day periods -- USE THIS FOR 32-DAY PERIODS

newshortfall2020 <- historic_data[historic_data$date >= "2020-08-31" & historic_data$date <= "2020-10-02", ]
newshortspring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-02-26", ]
newshortfall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-10-01", ]
newshortspring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-02-25", ]
newfall2022  <- historic_data[historic_data$date >= "2022-08-29",]

#devon 30 day periods - USE THIS FOR 30-day periods

d_newshortfall2020 <- historic_data[historic_data$date >= "2020-08-31" & historic_data$date <= "2020-09-30", ]
d_newshortspring2021 <- historic_data[historic_data$date >= "2021-01-25" & historic_data$date <= "2021-02-24", ]
d_newshortfall2021 <- historic_data[historic_data$date >= "2021-08-30" & historic_data$date <= "2021-09-29", ]
d_newshortspring2022  <- historic_data[historic_data$date >= "2022-01-24" & historic_data$date <= "2022-02-23", ]
d_newthirtyfall2022 <- historic_data[historic_data$date >= "2022-08-29" & historic_data$date <= "2022-09-28",]
d_newfall2022  <- historic_data[historic_data$date >= "2022-08-29",] #32 days in this one

d_newfall2023 <- historic_data[historic_data$date >= "2023-8-28",]

d_tryfall2022 <- historic_data[historic_data$date >= "2022-8-28" & historic_data$date <= "2022-09-05",]


```



```{r}
#ignore this
sum(newshortfall2020$count)
sum(newshortspring2021$count)
sum(newshortfall2021$count)
sum(newshortspring2022$count)
sum(newfall2022$count)

#devon sums with adjusted semesters -- ignore this
sum(d_shortfall2020$count)
sum(d_shortspring2021$count)
sum(d_shortfall2021$count)
sum(d_shortspring2022$count)
sum(d_fall2022$count)

#rina sums with new time periods -- USE THIS FOR 32-DAY COUNTS
sum(newshortfall2020$count)
sum(newshortspring2021$count)
sum(newshortfall2021$count)
sum(newshortspring2022$count)
sum(newfall2022$count)

#devon sums with 30-day periods -- USE THIS FOR 30-DAY COUNTS
sum(d_newshortfall2020$count)
sum(d_newshortspring2021$count)
sum(d_newshortfall2021$count)
sum(d_newshortspring2022$count)
sum(d_newthirtyfall2022$count)
sum(d_newfall2022$count)
sum(d_newfall2023$count)
sum(d_tryfall2022$count)
```



```{r}
historic_databydate <- historic_data %>% group_by(date)
historic_databydate <- subset(historic_databydate, select = -c(role))


newshortfall2020sr <-
  newshortfall2020%>%
  filter(type == "Self Reported")

newshortspring2021sr <-
  newshortspring2021%>%
  filter(type == "Self Reported")

newshortfall2021sr <-
  newshortfall2021%>%
  filter(type == "Self Reported")

newshortspring2022sr <-
  newshortspring2022%>%
  filter(type == "Self Reported")

newfall2022sr <-
  newfall2022%>%
  filter(type == "Self Reported")



```

```{r}
sum(newshortfall2020sr$count)
sum(newshortspring2021sr$count)
sum(newshortfall2021sr$count)
sum(newshortspring2022sr$count)
sum(newfall2022sr$count)



```


```{r}
sum(fall2020sr$count)
sum(spring2021sr$count)
sum(fall2021sr$count)
sum(spring2022sr$count)
sum(fall2022sr$count)
```

```{r}

newshortfall2020u <-
  newshortfall2020%>%
  filter(type =="UHC Reported")

shortspring2021u <-
  shortspring2021%>%
  filter(type == "UHC Reported")

shortfall2021u <-
  shortfall2021%>%
  filter(type == "UHC Reported")

shortspring2022u <-
  shortspring2022%>%
  filter(type == "UHC Reported")

fall2022u <-
  fall2022%>%
  filter(type == "UHC Reported")

```

```{r}
sum(fall2020u$count)
sum(spring2021u$count)
sum(fall2021u$count)
sum(spring2022u$count)
sum(fall2022u$count)
```


```{r}
sum(fall2020u$count) + sum(fall2020sr$count)
sum(shortspring2021u$count) + sum(shortspring2021sr$count)
sum(shortfall2021u$count)+sum(shortfall2021sr$count)
sum(shortspring2022u$count)+sum(shortspring2022sr$count)

sum(fall2022u$count)+sum(fall2022sr$count)

```

Pivoting, binding, and cleaning historic_data here -- Shreya --- old and irrelevant :wave:
```{r}

#load in historic_data
historic_data <- read_csv('historic_data.csv') %>%
  clean_names() %>%
  rename(
    date = positive_test_result_date
  ) %>%
  drop_na(question_from_laura) #%>%
  # mutate(
  #   number_of_cases = as.double(number_of_cases)
  # )

#clean date
historic_data$date <- as.Date(parse_date_time(historic_data$date, "dmy"))
historic_data$year <- as.numeric(format(historic_data$date, "%Y"))

clean_historic_historic_data <- historic_data %>%
  pivot_wider(names_from=primary_role, values_from=number_of_cases, values_fn = length, values_fill = 0) %>%
  clean_names() %>%
  mutate(
    student = as.character(student)
  )
  #mutate(across(student:faculty, ~replace(., lengths(.) == 0, NA)))

#export to excel because efficiency and I don't know how
write_xlsx(clean_historic_historic_data, "clean_historic_historic_data.xlsx")

self_reports_clean <- clean_historic_historic_data %>%
  filter(self_reported_or_uhc_reported == "Self Reported")

# test
fall2020 <- clean_historic_historic_data[clean_historic_historic_data$date >= "2020-08-30" & clean_historic_historic_data$date <= "2020-12-22", ]
self_reports_clean[is.null(self_reports_clean)] = 0


uhc_reports_clean <- clean_historic_historic_data %>%
  filter(self_reported_or_uhc_reported == "UHC Reported")

```


Joining the important policy dates to cases -- Shreya
```{r}

#load in historic_data
new_historic_data <- read_xlsx("latest_9_30_covid_historic_data.xlsx") %>%
  clean_names() %>%
  rename(
    date = positive_test_result_date
  )


### policy dates
historic_dates <- read_csv("historical_covid_dates.csv") %>%
  select(dates, what_happened) %>%
  rename(
    date = dates
  )


##pivoted historic_data
join_dates_and_cases <- new_historic_data %>%
  left_join(historic_dates, by=c("date")) %>%
  pivot_wider(names_from=primary_role, values_from=no_of_cases) %>%
  clean_names()

##wider for fun
wide_join_dates_and_cases <- new_historic_data %>%
  left_join(historic_dates, by=c("date")) %>%
  select(-diamondback_request) %>%
  group_by(reporting_type) %>%
  pivot_wider(names_from=c("primary_role", "reporting_type"), values_from=no_of_cases) %>%
  clean_names() %>%
  mutate_at(c(3:8), ~replace_na(.,0))


write_xlsx(join_dates_and_cases, "join_dates_and_cases.xlsx")

write_xlsx(wide_join_dates_and_cases, "wide_join_dates_and_cases.xlsx")



```
